---
title: "Proof of Concept: Aquatic Species-at-Risk"
subtitle: "by Natural Resource Region and District"
author: "Aquatic Ecosystems Branch"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute
---

```{css, echo=FALSE}

.Wrap {
  max-width: 100%;
}

.Content {
  padding: 0 5px 0 10px;
}

.leaflet-popup-content {
  width: 400px !important;
}

```

```{r setup, include = F}
# <style type="text/css">
# .page-inner {
#   padding: 10px 10px 20px 0px;
# }
# .book-body.fixed{
# margin-left: 150px;
# }
# </style>
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = F)

library(sf)
library(tidyverse)
library(leaflet)
```

```{r load_data}
nr_r = bcmaps::nr_regions() |> 
  sf::st_transform(crs = 4326) |> 
  rmapshaper::ms_simplify()

nr_d = bcmaps::nr_districts() |> 
  sf::st_transform(crs = 4326) |> 
  rmapshaper::ms_simplify()

sar = sf::read_sf('app/www/species_ecosystems_at_risk_publically_available.gpkg') |> 
  dplyr::mutate(TAX_CLASS = stringr::str_to_title(TAX_CLASS))

unique_taxes = unique(sar$TAX_CLASS)
unique_taxes = unique_taxes[order(unique_taxes)]

wilson_sar = readr::read_csv('app/www/Aquatic_SAR_GregWilson.csv')
```

## Basic Data Description

Using the Conservation Data Centre's <a href = 'https://catalogue.data.gov.bc.ca/dataset/species-and-ecosystems-at-risk-publicly-available-occurrences-cdc'>publically available species- and ecosystems-at-risk layer</a>, there are `r nrow(sar)` rows of SAR data if we include the following taxonomic classes: `r unique_taxes`

```{r filter_sar}
sar_f = sar |> 
  dplyr::filter(TAX_CLASS %in% c('Amphibians','Bivalves','Lampreys','Ray-Finned Fishes')) 
```

After filtering for just Amphibians, Bivalves, Lampreys, and Ray-Finned Fishes, there are `r nrow(sar_f)` rows.

```{r reproject_sar}
sar_f = sf::st_transform(sar_f, 4326)
```

```{r sum_sar_by_region_and_district}
sar_by_district = sar_f |> 
  sf::st_join(nr_d |> dplyr::select(District = DISTRICT_NAME), 
              st_intersects) |> 
  st_drop_geometry() |> 
  count(ENG_NAME, District) |> 
  group_by(District) |> 
  reframe(all_info = paste0(ENG_NAME, " (",n,")"),
          total_n = sum(n)) |> 
  group_by(District) |> 
  summarise(all_info = paste0(all_info, collapse = ', '),
            total_n = mean(total_n))
  
sar_by_region = sar_f |> 
  sf::st_join(nr_r |> dplyr::select(Region = REGION_NAME), st_intersects) |> 
  st_drop_geometry() |> 
  count(ENG_NAME, Region) |> 
  group_by(Region) |> 
  reframe(all_info = paste0(ENG_NAME, " (",n,")"),
          total_n = sum(n)) |> 
  group_by(Region) |> 
  summarise(all_info = paste0(all_info, collapse = ', '),
            total_n = mean(total_n))
```

```{r add_greg_excel_data}
wilson_sar_to_join = wilson_sar |> 
  dplyr::mutate(nrs_region = dplyr::case_when(
    str_detect(nrs_region, 'west') ~ 'West Coast Natural Resource Region',
    str_detect(nrs_region, 'south') ~ 'South Coast Natural Resource Region',
    str_detect(nrs_region, 'Thompson') ~ 'Thompson-Okanagan Natural Resource Region',
    str_detect(nrs_region, 'Kootenay') ~ 'Kootenay-Boundary Natural Resource Region',
    str_detect(nrs_region, 'Cariboo') ~ 'Cariboo Natural Resource Region',
    str_detect(nrs_region, 'Omineca') ~ 'Omineca Natural Resource Region',
    T ~ 'Not matched!'
  )) |> 
  dplyr::rename(Region = nrs_region) |> 
  group_by(Region) |> 
  summarise(across(everything(), \(x) paste0(x, collapse = '; '))) |> 
  dplyr::mutate(across(status:key_areas_of_concern, \(x) str_remove_all(x, 'NA(; )?'))) |> 
  dplyr::mutate(across(status:key_areas_of_concern, \(x) str_to_sentence(x)))

sar_by_region = sar_by_region |> 
  dplyr::left_join(wilson_sar_to_join)

# Spatial match of regions and districts so that we can port Greg Wilson's excel data 
# from the region scale to the districts (although it will be repeated.)

district_region_lookup_table = nr_d |> 
  dplyr::select(Region = REGION_ORG_UNIT_NAME,
                District = DISTRICT_NAME) |> 
  sf::st_drop_geometry()

sar_by_district = sar_by_district |> 
  dplyr::left_join(
    wilson_sar_to_join |> 
      left_join(district_region_lookup_table) |> 
      dplyr::select(-Region)
  )
```

## Interactive Map of Natural Resource Regions / Districts

*Toggle between regions and districts by clicking the radio-buttons at the bottom-right!* *Click on the map shapes to access more detail*

```{r organize_regions_and_districts_alphabetically}
nr_r = nr_r |> 
  dplyr::arrange(REGION_NAME)

nr_d = nr_d |> 
  dplyr::arrange(DISTRICT_NAME)
```

```{r make_leaflet_tables}
district_table = leafpop::popupTable(
  sar_by_district |> 
    mutate(District = stringr::str_remove_all(District,
                                              " Natural.*")) |> 
    left_join(
      sar_by_district |> 
        mutate(District = stringr::str_remove_all(District,
                                                " Natural.*")) |> 
        tidyr::separate_longer_delim(all_info, delim = ', ') |> 
        dplyr::count(District, name = 'distinct_SAR') |> 
        dplyr::mutate(distinct_SAR = replace_na(distinct_SAR, 0))
    ) |> 
    dplyr::select(
      `BCDC SAR` = all_info,
      `BCDC # Records` = total_n,
      `BCDC # Species` = distinct_SAR,
      `Expert-ID SAR` = species_common_name_population,
      Status = status,
      `Major Watershed(s)` = major_watershed,
      Threats = threats,
      `Key Areas of Concern` = key_areas_of_concern
    ),
  className = 'table_wide'
)

region_table = leafpop::popupTable(
  sar_by_region |> 
    mutate(Region = stringr::str_remove_all(Region,
                                              " Natural.*")) |> 
    left_join(
      sar_by_region |> 
        mutate(Region = stringr::str_remove_all(Region,
                                                " Natural.*")) |> 
        tidyr::separate_longer_delim(all_info, delim = ', ') |> 
        dplyr::count(Region, name = 'distinct_SAR') |> 
        dplyr::mutate(distinct_SAR = replace_na(distinct_SAR, 0))
    ) |> 
    dplyr::select(
      `BCDC SAR` = all_info,
      `BCDC # Records` = total_n,
      `BCDC # Species` = distinct_SAR,
      `Expert-ID SAR` = species_common_name_population,
      Status = status,
      `Major Watershed(s)` = major_watershed,
      Threats = threats,
      `Key Areas of Concern` = key_areas_of_concern
    ),
  className = 'table_wide'
)
```

```{r make_leaflet_colour_palettes}
sar_col_pal = leaflet::colorFactor(
  palette = 'Spectral',
  domain = sar_f$TAX_CLASS
)

nr_r = nr_r |> 
  left_join(
    sar_by_region |> 
  tidyr::separate_longer_delim(all_info, delim = ', ') |> 
  dplyr::count(Region, name = 'distinct_SAR') |> 
    dplyr::rename(REGION_NAME = Region)
  )

nr_d = nr_d |> 
  left_join(
    sar_by_district |> 
  tidyr::separate_longer_delim(all_info, delim = ', ') |> 
  dplyr::count(District, name = 'distinct_SAR') |> 
    dplyr::rename(DISTRICT_NAME = District)
  ) |> 
  dplyr::mutate(distinct_SAR = replace_na(distinct_SAR, 0))

region_colour_palette = colorNumeric(
  palette = 'Purples',
  domain = nr_r$distinct_SAR
)

district_colour_palette = colorNumeric(
  palette = 'Purples',
  domain = nr_d$distinct_SAR
)
```

```{r leaflet_map, fig.width=12, fig.align='center', fig.height = 7}
leaflet() |> 
  addTiles(group = 'OSM') |> 
  addProviderTiles(provider = providers$CartoDB,
                   group = 'cartoDB') |> 
  addPolygons(
    color = 'black',
    fillColor = ~region_colour_palette(distinct_SAR),
    fillOpacity = 0.6,
    label = ~paste0(REGION_NAME,": ",distinct_SAR," distinct SAR"),
    popup = ~lapply(region_table, \(x) htmltools::HTML(x[!is.na(x) | x != ""])),
    group = 'Regions',
    weight = 2,
    data = nr_r) |>
  addPolygons(
    color = 'black',
    fillColor = ~district_colour_palette(distinct_SAR),
    fillOpacity = 0.6,
    label = ~paste0(DISTRICT_NAME,": ",distinct_SAR," distinct Aquatic SAR"),
    popup = ~lapply(district_table, \(x) htmltools::HTML(x[!is.na(x) | x != ""])),
    group = 'Districts',
    weight = 2,
    data = nr_d) |> 
  addPolygons(
    label = ~ENG_NAME,
    color = ~sar_col_pal(TAX_CLASS),
    fillColor = ~sar_col_pal(TAX_CLASS),
    fillOpacity = 1,
    weight = 2,
    data = sar_f,
    group = 'Aquatic SAR'
  ) |> 
  addLegend(
    pal = sar_col_pal,
    values = sar_f$TAX_CLASS
  ) |> 
  addLegend(
    title = 'Number of \nDistinct \nAquatic SAR',
    colors = c('white','#ad80cf','purple'),
    values = c(1,9,18),
    labels = c('few','moderate','many')
  ) |> 
  addLayersControl('bottomright',
                   baseGroups = c('Regions','Districts'),
                   overlayGroups = c('Aquatic SAR'),
                   options = layersControlOptions(collapsed = F)) |> 
  leaflet::hideGroup('nr_d') |> 
  addScaleBar('bottomright') |> 
  leaflet.extras::addResetMapButton() |> 
  htmlwidgets::onRender("
    function(el, x) {
      this.on('baselayerchange', function(e) {
        e.layer.bringToBack();
      })
    }
  ")
```

## Interactive Tables {.tabset}

```{r}
library(DT)
```

### Region

```{r data_table_region}
region_table_dat = sar_by_region |> 
    mutate(Region = stringr::str_remove_all(Region,
                                              " Natural.*")) |> 
    left_join(
      sar_by_region |> 
        mutate(Region = stringr::str_remove_all(Region,
                                                " Natural.*")) |> 
        tidyr::separate_longer_delim(all_info, delim = ', ') |> 
        dplyr::count(Region, name = 'distinct_SAR') |> 
        dplyr::mutate(distinct_SAR = replace_na(distinct_SAR, 0))
    ) |>
    dplyr::select(
      Region,
      `BCDC SAR` = all_info,
      `BCDC # Records` = total_n,
      `BCDC # Species` = distinct_SAR,
      `Expert-ID SAR` = species_common_name_population,
      Status = status,
      `Major Watershed(s)` = major_watershed,
      Threats = threats,
      `Key Areas of Concern` = key_areas_of_concern
    )

DT::datatable(
  region_table_dat,
  options = list(
    paging = TRUE,
    pageLength = 5,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'font-size': '100%'});",
      "}"))) |> DT::formatStyle(columns = names(region_table_dat), 
                                fontSize = '100%',
      ) |> 
  DT::formatStyle(
    columns = c(5:9),
    backgroundColor = 'lightblue'
  )
```

Note:

1.  *BCDC = BC Data Catalogue*
2.  *Expert-ID SAR = Aquatic species-at-risk identified by provincial subject matter expert*
3.  *Grey cells in table are based on BCDC datasets, light blue cells based on subject matter expert*

### District

```{r data_table_district}
district_table_dat = sar_by_district |> 
    mutate(District = stringr::str_remove_all(District,
                                              " Natural.*")) |> 
    left_join(
      sar_by_district |> 
        mutate(District = stringr::str_remove_all(District,
                                                " Natural.*")) |> 
        tidyr::separate_longer_delim(all_info, delim = ', ') |> 
        dplyr::count(District, name = 'distinct_SAR') |> 
        dplyr::mutate(distinct_SAR = replace_na(distinct_SAR, 0))
    ) |> 
    dplyr::select(
      District,
      `BCDC SAR` = all_info,
      `BCDC # Records` = total_n,
      `BCDC # Species` = distinct_SAR,
      `Expert-ID SAR` = species_common_name_population,
      Status = status,
      `Major Watershed(s)` = major_watershed,
      Threats = threats,
      `Key Areas of Concern` = key_areas_of_concern
    )

DT::datatable(
  district_table_dat,
  options = list(
    paging = TRUE,
    pageLength = 5#,
    # initComplete = JS(
    #   "function(settings, json) {",
    #   "$(this.api().table().header()).css({'font-size': '100%'});
    #   $(this.api().table().header()[3]).css({'background:lightblue': '100%'});",
    #   "}")
    )
  ) |> DT::formatStyle(columns = names(district_table_dat), 
                                fontSize = '100%',
      ) |> 
  DT::formatStyle(
    columns = c(5:9),
    backgroundColor = 'lightblue'
  )
```
